\section*{Acknowledgments}


I would like to thank everyone who contributed to the success of my PhD project. 

Many thanks to my university supervisor Prof. Dr. Johannes Wessels for giving me the possibility of accomplishing my Master and PhD thesis at CERN, while being part of his working group at the University of M\"{u}nster. 

Special thanks to my CERN supervisor Dr. Roderik Bruce. Thank you very much for the intelligent structuring of the thesis work and for the many interesting and inspiring discussions!

It was a great experience to be part of the LHC collimation team and I would like to thank the whole team for making my three years at CERN such a pleasure: Roderik Bruce, Maria Fiascaris, Hector Garcia Morales, Regina Kwee-Hinzmann, Aurelien Marsili, Alessio Mereghetti, Daniele Mirarchi, Elena Quaranta, Stefano Redaelli, Adriana Rossi, Roberto Rossi, Belen Salvachua, Alessia Valloni, Gianluca Valentino and Joschka Wagner. Many thanks to Dr. Stefano Redaelli for welcoming me in his team and for his constant support during my PhD time. 

Thanks for constant support during my thesis work to John Jowett, Riccardo de Maria and Massimo Gionvannozzi.

My acknowledgments to the CERN FLUKA team for the help with the setup of the hiSixTrack-FLUKA coupling: F. Cerutti, A. Ferrari, A. Lechner, A. Mereghetti, P. G. Ortega, E. Skordis, V. Vlachoudis. 

Many thanks to my office mates: D. Mirarchi, E. Quaranta, T. Mertens, M. Schaumann. 

For proof-reading my thesis I would like to thank R. Bruce, B. Holzer, E. Sicking, A. Mereghetti and M. Breitenfeld.


But also outside of work some important people in my life contributed to this thesis, by giving me inspiration and by making my time outside work worthwhile. Most importantly, my wife Julia, with whom I shared all the happy and the hard times during the last three years. She was my haven of tranquility and graced my life with pleasure and ease. 

I thank my family Stefanie, Wolfgang, Vanessa, Johanna, Maria and Litwina for everything they did for me.

For interesting coffee breaks and activities outside CERN, I would like to thank Robert Mertzig, Andy Langner and Martin Breitenfeld.

I would like to thank Sam Ringle for our stimulating friendship and his constant presence in my life for over two decades. I am also very grateful to Alex Senzig, not only for his musical contributions, but also for his loyal friendship which enriched my life for many years now. 

Last but not least I would like to thank Andy, Robert, Martin, Eva Dietz, Eva Sicking, Vanessa, Georg, Desi, Gabriele, Werner and many more.